---
title: "Extra Credit Assignment Case 01"
author: "Sara Lemus"
date: "2/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(survival)
library(survminer)
library(DescTools)
library(broom)
```

```{r}
data <- read_csv("dig.csv")
```

## Table 1

```{r}
#renaming levels for display purposes
data <- data %>% 
  mutate(TRTMT = case_when(TRTMT == 0 ~ "Placebo",
                                TRTMT == 1 ~ "Digoxin"))
data <- data %>% 
  mutate(TRTMT = as.factor(TRTMT))
```

Age- Mean and SD
```{r}
# mean and sd age
data %>% 
  group_by(TRTMT) %>% 
  summarise(mean_age = round(mean(AGE),1)) 
data %>% 
  group_by(TRTMT) %>% 
  summarise(sd_age = round(sd(AGE),1))
```

Ejection fraction- mean and SD
```{r}
# ejection fraction
data %>% 
  group_by(TRTMT) %>% 
  summarise(mean_age = round(mean(EJF_PER),1)) 
data %>% 
  group_by(TRTMT) %>% 
  summarise(sd_age = round(sd(EJF_PER),1))
```
Median duration of CHF- Mo
```{r}
data %>% 
  group_by(TRTMT) %>% 
  filter(!is.na(CHFDUR)) %>% 
  summarise(median = median(CHFDUR))
```

Female, Nonwhite, Age > 70
```{r}
#female
data %>% 
  group_by(TRTMT) %>% 
  count(SEX) %>% 
  mutate(prop_female = round(n/sum(n) * 100,1)) %>% 
  filter(SEX == 2) %>% 
  select(prop_female)

#nonwhite
data %>% 
  group_by(TRTMT) %>% 
  count(RACE) %>% 
  mutate(prop_nonwhite = round(n/sum(n) * 100,1)) %>% 
  filter(RACE == 2) %>% 
  select(prop_nonwhite)

#age > 70
data %>% 
  group_by(TRTMT) %>%
  count(AGE > 70) %>% 
  mutate(prop_70plus = round(n/sum(n) * 100,1))%>% 
  filter(TRUE) %>% 
  filter(`AGE > 70` == "TRUE") %>% 
  select(prop_70plus)
```

Method of assesing ejection fraction, Cardiothotacic ratio
```{r}
#ejf meth
data %>% 
  group_by(TRTMT) %>%
  mutate(EJFMETH = case_when(EJFMETH == 1 ~ "Radionuclide",
                             EJFMETH == 2 ~ "Angiography",
                             EJFMETH == 3 ~ "2-D Echo")) %>% 
  count(EJFMETH) %>% 
  mutate(prop_ejfmeth = round(n/sum(n) * 100,1)) 

#cardiothotacic ratio >0.55
data %>% 
  group_by(TRTMT) %>%
  count(CHESTX > 0.55) %>% 
  mutate(prop_chest = round(n/sum(n) * 100,1)) %>% 
  filter(`CHESTX > 0.55` == "TRUE") %>% 
  select(prop_chest)
```
NYHA class, No. of Signs or symptoms of CHF
```{r}
#CHF
data %>% 
  group_by(TRTMT) %>%
  filter(!is.na(FUNCTCLS)) %>% 
  mutate(FUNCTCLS = case_when(FUNCTCLS == 1 ~ "I",
                             FUNCTCLS == 2 ~ "II",
                             FUNCTCLS == 3 ~ "III",
                             FUNCTCLS == 4 ~ "IV")) %>% 
  count(FUNCTCLS) %>% 
  mutate(prop_nyha = round(n/sum(n) * 100,1)) 
#no of symptoms of CHF
data %>% 
  group_by(TRTMT) %>%
  count(NSYM) %>% 
  mutate(prop_chp = round(n/sum(n) * 100,1)) %>% 
  select(NSYM, prop_chp)
```

Medical History
```{r}
#medical history 
data %>% 
  group_by(TRTMT) %>%
  count(PREVMI) %>% 
  mutate(prop_prevmi = round(n/sum(n) * 100,1)) %>% 
  filter(PREVMI == 1) %>% 
  select(prop_prevmi)
data %>% 
  group_by(TRTMT) %>%
  count(ANGINA) %>% 
  mutate(prop_angina = round(n/sum(n) * 100,1)) %>% 
  filter(ANGINA == 1) %>% 
  select(prop_angina)
data %>% 
  group_by(TRTMT) %>%
  count(DIABETES) %>% 
  mutate(prop_dia = round(n/sum(n) * 100,1)) %>% 
  filter(DIABETES == 1) %>% 
  select(prop_dia)
data %>% 
  group_by(TRTMT) %>%
  count(HYPERTEN) %>% 
  mutate(prop_hyper = round(n/sum(n) * 100,1)) %>% 
  filter(HYPERTEN == 1) %>% 
  select(prop_hyper)
```

Previous digoxin use
```{r}
#prev digoxin
data %>% 
  group_by(TRTMT) %>% 
  count(DIGUSE) %>% 
  mutate(prop_dig = round(n/sum(n) * 100,1)) %>% 
  filter(DIGUSE == 1) %>% 
  select(prop_dig)
```
Primary cause of CHF
```{r}
data %>% 
  #filter(CHFETIOL != 6) %>%
  mutate(CHFETIOL = case_when(CHFETIOL == 1 ~ "Ischemic", 
                              is.na(CHFETIOL) ~ "null",
                              TRUE ~ "Nonischemic")) %>% 
  group_by(TRTMT) %>% 
  count(CHFETIOL) %>% 
  mutate(prop_chfcause = round(n/sum(n) * 100, 1)) %>% 
  filter(CHFETIOL != "null") 
```
```{r}
data %>% 
  mutate(CHFETIOL = case_when(CHFETIOL == 1 ~ "Ischemic",
                              CHFETIOL == 4 ~ "Idiopathic", 
                              CHFETIOL == 2 ~ "Hypertensive",
                              CHFETIOL == 3 | CHFETIOL == 5 | CHFETIOL == 6 ~ "Other")) %>% 
  group_by(TRTMT) %>% 
  count(CHFETIOL) %>% 
  mutate(prop_chfcause = round(n/sum(n) * 100,1)) %>% 
  filter(!is.na(CHFETIOL) & CHFETIOL != "Ischemic") 
```

Concomitant medications
```{r}
#concomitant meds
data <- data %>% 
  mutate(diuretic_used = case_when(DIURET == 1 ~ 1,
                                   DIURETK == 1 ~ 1,
                                   TRUE ~ 0))
data %>% 
  group_by(TRTMT) %>% 
  count(diuretic_used) %>% 
  mutate(prop_diuretics = round(n/sum(n) * 100,1)) %>% 
  filter(diuretic_used == 1) %>% 
  select(prop_diuretics)
data %>% 
  group_by(TRTMT) %>% 
  count(ACEINHIB) %>% 
  mutate(prop_acei = round(n/sum(n) * 100,1)) %>% 
  filter(ACEINHIB == 1) %>% 
  select(prop_acei)
#off by .1
data %>% 
  group_by(TRTMT) %>% 
  count(NITRATES) %>% 
  mutate(prop_nitr = round(round(n/sum(n) * 100,2),1)) %>% 
  filter(NITRATES == 1) %>% 
  select(prop_nitr)
data %>% 
  group_by(TRTMT) %>% 
  count(VASOD) %>% 
  mutate(prop_vasod = round(n/sum(n) * 100,1)) %>% 
  filter(VASOD == 1) %>% 
  select(prop_vasod)
```

Daily dose of study medication perscribed
```{r}
#daily dose
data %>% 
  group_by(TRTMT) %>% 
  count(DIGDOSE) %>% 
  mutate(prop_dose = round(n/sum(n) * 100,1)) %>% 
  select(DIGDOSE, prop_dose) %>% 
  filter(DIGDOSE == 0.125 | DIGDOSE == 0.250 | DIGDOSE == 0.375 | DIGDOSE == 0.500)
```


## Table 4
```{r}
#relevel
data <- data %>% 
  mutate(TRTMT_RE = fct_relevel(TRTMT, "Placebo")) 
```

### Ejection Fraction
```{r}
data <- data %>% 
  mutate(EJF_PER_NEW = case_when(EJF_PER >= 25  ~ "0.25-0.45",
                                  TRUE ~ "<0.25"))
```

```{r}
#ejection fraction
table_1_ejf <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(EJF_PER_NEW) 
table_2_ejf <- data %>% 
  group_by(TRTMT) %>% 
  count(EJF_PER_NEW) 
table_1_ejf <- table_1_ejf %>% 
  rename("yes_dwhf" = n)
table_2_ejf <- table_2_ejf %>% 
  rename("randomized" = n)
result_table_ejf <- inner_join(table_1_ejf, table_2_ejf) 
```

```{r}
result_table_ejf <- result_table_ejf %>% 
  mutate(percent = yes_dwhf / randomized) 
result_table_ejf %>% 
  mutate(percent = round(yes_dwhf / randomized, 3)* 100)
```

Absolute difference
```{r}
#https://stackoverflow.com/questions/7381455/filtering-a-data-frame-by-values-in-a-column
yes_result_1_ejf <- result_table_ejf$yes_dwhf[result_table_ejf$EJF_PER_NEW == "0.25-0.45"]
no_result_1_ejf <- result_table_ejf$randomized[result_table_ejf$EJF_PER_NEW == "0.25-0.45"]

#https://rdrr.io/cran/DescTools/man/BinomDiffCI.html
round(BinomDiffCI(yes_result_1_ejf[1], no_result_1_ejf[1], yes_result_1_ejf[2], no_result_1_ejf[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

```{r}
yes_result_0_ejf <- result_table_ejf$yes_dwhf[result_table_ejf$EJF_PER_NEW == "<0.25"]
no_result_0_ejf <- result_table_ejf$randomized[result_table_ejf$EJF_PER_NEW == "<0.25"]

round(BinomDiffCI(yes_result_0_ejf[1], no_result_0_ejf[1], yes_result_0_ejf[2], no_result_0_ejf[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio 
```{r}
#ejf fraction
data1 <- data %>% 
  filter(EJF_PER_NEW == "0.25-0.45")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data1))$conf.int, 2)

data2 <- data %>% 
  filter(EJF_PER_NEW == "<0.25")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data2))$conf.int, 2)
```

### Previous Use of Digoxin

```{r}
#prev dig use
table_1 <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(DIGUSE) 
table_2 <- data %>% 
  group_by(TRTMT) %>% 
  count(DIGUSE) 
table_1 <- table_1 %>% 
  rename("yes_dwhf" = n)
table_2 <- table_2 %>% 
  rename("randomized" = n)
result_table <- inner_join(table_1, table_2) 
```

```{r}
result_table <- result_table %>% 
  mutate(percent =round(yes_dwhf / randomized, 3)* 100) 
result_table 
```

Absolute difference
```{r}
#https://stackoverflow.com/questions/7381455/filtering-a-data-frame-by-values-in-a-column
yes_result_1 <- result_table$yes_dwhf[result_table$DIGUSE == 1]
no_result_1 <- result_table$randomized[result_table$DIGUSE == 1]

#https://rdrr.io/cran/DescTools/man/BinomDiffCI.html
round(BinomDiffCI(yes_result_1[1], no_result_1[1], yes_result_1[2], no_result_1[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

```{r}
yes_result_0 <- result_table$yes_dwhf[result_table$DIGUSE == 0]
no_result_0 <- result_table$randomized[result_table$DIGUSE == 0]

round(BinomDiffCI(yes_result_0[1], no_result_0[1], yes_result_0[2], no_result_0[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio
```{r}
#prev dig use
data4 <- data %>% 
  filter(DIGUSE == 1)
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data4))$conf.int, 2)

data3 <- data %>% 
  filter(DIGUSE == 0)
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data3))$conf.int, 2)
```

### Cause of Heart Failure

```{r}
data <- data %>% 
  mutate(CHFETIOL_NEW = case_when(CHFETIOL == 1 ~ "Ischemic", 
                              is.na(CHFETIOL) ~ "null",
                              TRUE ~ "Nonischemic"))
```

```{r}
#hart failure
table_1_cause <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(CHFETIOL_NEW) %>% 
  filter(CHFETIOL_NEW != "null") 
table_2_cause <- data %>% 
  group_by(TRTMT) %>% 
  count(CHFETIOL_NEW) %>% 
  filter(CHFETIOL_NEW != "null") 
table_1_cause <- table_1_cause %>% 
  rename("yes_dwhf" = n)
table_2_cause <- table_2_cause %>% 
  rename("randomized" = n)
result_table_cause <- inner_join(table_1_cause, table_2_cause) 
```

```{r}
result_table_cause <- result_table_cause %>% 
  mutate(percent = yes_dwhf / randomized) 
result_table_cause %>% 
  mutate(percent = round(yes_dwhf / randomized, 3)* 100)
```

Absolute difference
```{r}
#https://stackoverflow.com/questions/7381455/filtering-a-data-frame-by-values-in-a-column
yes_result_1_cause <- result_table_cause$yes_dwhf[result_table_cause$CHFETIOL_NEW == "Ischemic"]
no_result_1_cause <- result_table_cause$randomized[result_table_cause$CHFETIOL_NEW == "Ischemic"]

#https://rdrr.io/cran/DescTools/man/BinomDiffCI.html
round(BinomDiffCI(yes_result_1_cause[1], no_result_1_cause[1], yes_result_1_cause[2], no_result_1_cause[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

```{r}
yes_result_0_cause <- result_table_cause$yes_dwhf[result_table_cause$CHFETIOL_NEW == "Nonischemic"]
no_result_0_cause <- result_table_cause$randomized[result_table_cause$CHFETIOL_NEW == "Nonischemic"]

round(BinomDiffCI(yes_result_0_cause[1], no_result_0_cause[1], yes_result_0_cause[2], no_result_0_cause[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio 
```{r}
#cause of heart failure
data5 <- data %>% 
  filter(CHFETIOL_NEW == "Ischemic")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data5))$conf.int, 2)

data6 <- data %>% 
  filter(CHFETIOL_NEW == "Nonischemic")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data6))$conf.int, 2)
```



### Cardiothoracic Ratio

```{r}
data <- data %>% 
  mutate(CHESTX_NEW = case_when(CHESTX > 0.55 ~ ">0.55", 
                                TRUE ~ "<0.55"))
```

```{r}
table_1_chest <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(CHESTX_NEW)
table_2_chest <- data %>% 
  group_by(TRTMT) %>% 
  count(CHESTX_NEW)
table_1_chest <- table_1_chest %>% 
  rename("yes_dwhf" = n)
table_2_chest <- table_2_chest %>% 
  rename("randomized" = n)
result_table_chest <- inner_join(table_1_chest, table_2_chest) 
```

```{r}
result_table_chest <- result_table_chest %>% 
  mutate(percent = yes_dwhf / randomized) 
result_table_chest %>% 
  mutate(percent = round(yes_dwhf / randomized, 3)* 100)
```

Absolute difference
```{r}
#https://stackoverflow.com/questions/7381455/filtering-a-data-frame-by-values-in-a-column
yes_result_1_chest <- result_table_chest$yes_dwhf[result_table_chest$CHESTX_NEW == "<0.55"]
no_result_1_chest <- result_table_chest$randomized[result_table_chest$CHESTX_NEW == "<0.55"]

#https://rdrr.io/cran/DescTools/man/BinomDiffCI.html
round(BinomDiffCI(yes_result_1_chest[1], no_result_1_chest[1], yes_result_1_chest[2], no_result_1_chest[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```


```{r}
yes_result_0_chest <- result_table_chest$yes_dwhf[result_table_chest$CHESTX_NEW == ">0.55"]
no_result_0_chest <- result_table_chest$randomized[result_table_chest$CHESTX_NEW == ">0.55"]

round(BinomDiffCI(yes_result_0_chest[1], no_result_0_chest[1], yes_result_0_chest[2], no_result_0_chest[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio 
```{r}
#cardiothoracic ratio
data7 <- data %>% 
  filter(CHESTX_NEW == "<0.55")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data7))$conf.int, 2)

data8 <- data %>% 
  filter(CHESTX_NEW == ">0.55")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data8))$conf.int, 2)
```

### NYHA Class

```{r}
data <- data %>% 
  mutate(NYHA_NEW = case_when(FUNCTCLS == 1 | FUNCTCLS == 2 ~ "I or II", 
                              TRUE ~ "III or IV"))
```

```{r}
#NYHA class
table_1_nyha <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(NYHA_NEW)
table_2_nyha <- data %>% 
  group_by(TRTMT) %>% 
  count(NYHA_NEW)
table_1_nyha <- table_1_nyha %>% 
  rename("yes_dwhf" = n)
table_2_nyha <- table_2_nyha %>% 
  rename("randomized" = n)
result_table_nyha <- inner_join(table_1_nyha, table_2_nyha) 
```

```{r}
result_table_nyha <- result_table_nyha %>% 
  mutate(percent = yes_dwhf / randomized) 
result_table_nyha %>% 
  mutate(percent = round(yes_dwhf / randomized, 3)* 100)
```

Absolute difference
```{r}
#https://stackoverflow.com/questions/7381455/filtering-a-data-frame-by-values-in-a-column
yes_result_1_nyha <- result_table_nyha$yes_dwhf[result_table_nyha$NYHA_NEW == "I or II"]
no_result_1_nyha <- result_table_nyha$randomized[result_table_nyha$NYHA_NEW == "I or II"]

#https://rdrr.io/cran/DescTools/man/BinomDiffCI.html
round(BinomDiffCI(yes_result_1_nyha[1], no_result_1_nyha[1], yes_result_1_nyha[2], no_result_1_nyha[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

```{r}
#issue
yes_result_0_nyha <- result_table_nyha$yes_dwhf[result_table_nyha$NYHA_NEW== "III or IV"]
no_result_0_nyha <- result_table_nyha$randomized[result_table_nyha$NYHA_NEW== "III or IV"]

round(BinomDiffCI(yes_result_0_nyha[1], no_result_0_nyha[1], yes_result_0_nyha[2], no_result_0_nyha[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio 
```{r}
#nyha class
data9 <- data %>% 
  filter(NYHA_NEW == "I or II")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data9))$conf.int, 2)

#lower .95 off by 0.1
data10 <- data %>% 
  filter(NYHA_NEW == "III or IV")
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data10))$conf.int, 2)
```

### Overall study population

```{r}
data %>% 
  group_by(TRTMT) %>%
  count(DWHF) %>% 
  mutate(prop_dwhf = round(n/sum(n) * 100,1)) %>% 
  filter(DWHF == 1) %>% 
  select(prop_dwhf)
```

```{r}
result_table_trtmt <- data %>% 
  group_by(TRTMT) %>%
  count(DWHF) %>% 
  mutate(prop_dwhf = n/sum(n) * 100)
#result_table_trtmt %>% 
  #mutate(prop_dwhf = round(prop_dwhf, 1))
```

```{r}
table_1_trtmt <- data %>% 
  group_by(TRTMT) %>% 
  filter(DWHF == 1) %>% 
  count(DWHF)
table_2_trtmt <- data %>% 
  count(TRTMT) 
table_1_trtmt <- table_1_trtmt %>% 
  rename("yes_dwhf" = n)
table_2_trtmt <- table_2_trtmt %>% 
  rename("randomized" = n)
result_table_trtmt <- inner_join(table_1_trtmt, table_2_trtmt) 
```

Absolute Difference
```{r}
round(BinomDiffCI(result_table_trtmt$yes_dwhf[1], result_table_trtmt$randomized[1],result_table_trtmt$yes_dwhf[2], result_table_trtmt$randomized[2],
        conf.level = .95,
        method = "wald"), 3) * 100
```

Risk Ratio
```{r}
#study pop
round(summary(coxph(Surv(DWHFDAYS, DWHF) ~ TRTMT_RE, data))$conf.int, 2)
```



## Remarks
In general, I was able to replicate the findings of this study. However, there were some small discrepancies. In Table 1, 9 table values were off by about 0.1. This is assumed to be due to discrepancies in rounding, as the values in the original table were rounded up to the nearest 0.1, though the actual proportion estimates were most likely calculated with more signficant digits. In Table 4, the largest discrepancy was that for the "III or IV" category of NYHA class, the proportion for the placebo group differentiated by 0.1 and the estimate for the absolute difference of the two proportions differentiated by 0.2. The accompanying 95% confidence interval is also different by a similar margin. Once again, due to the fact that the magnitude of the error is small and that the other level of the variable (class I or II) mirrors the original table, it is also thought that this discrepancy may be also due to rounding, despite the fact that calculations were made follwing the paper (ie: absolute differences were calculated before rounding).










